---
title: "Aides Drug"
author: "Chris Gonzalez"
date: "June 14, 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```
```{r}
charlie<-filter(triUmp, DrugUser=="yes") 
```
```{r eval=FALSE}
charlie <- charlie %>%
mutate(across(where(is.character), as.factor))
```




# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

Clinical trials are a crucial step in determining the safety and efficacy of new drugs before they are approved for widespread use. In this particular scenario, a new preventive AIDS drug called triUmpsal is being tested in clinical trials. The focus of interest lies in evaluating the effectiveness of the drug specifically among the population of DrugUsers. This subset of individuals is of particular importance as they are more likely to benefit from preventive interventions due to their higher risk of HIV transmission.

To assess the effectiveness of the triUmpsal drug among DrugUsers, we can utilize the power of data analysis using tools like R and the dplyr package. By applying filtering techniques provided by dplyr, we can isolate the DrugUsers subset from the larger population and examine their response to the drug. This analysis aims to determine whether there is evidence to suggest that the triUmpsal drug is effective in reducing the incidence of HIV transmission among DrugUsers.

Through this investigation, we can gain insights into the potential impact of the triUmpsal drug on the specific target population and contribute valuable information to inform future decisions regarding its development and deployment. The findings may provide evidence for the effectiveness of the drug and guide further research or implementation strategies to combat the spread of HIV among DrugUsers, thereby offering hope for improved prevention measures and better healthcare outcomes.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The purpose of the present study is to investigate the potential efficacy of TriUmpsal as a drug for preventing AIDS. To facilitate this analysis, a box plot will be utilized as a visual tool to explore any potential relationship between the use of the drug TriUmpsal and a placebo as the control in preventing aids. Additionally, numerical data and inferential techniques will be employed to further analyze and interpret the findings, aiming to determine if TriUmpsal exhibits any correlation with AIDS prevention.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=charlie)
barchartGC(~infected + treatment,data=charlie, type="percent")

```
```{r}
library(ggplot2)

# Calculate percentage of each treatment group
percentage <- prop.table(table(charlie$treatment)) * 100

# Create a data frame with treatment and percentage
df_percentage <- data.frame(treatment = names(percentage), percentage = percentage)

# Bar chart for "infected" variable
chart_infected <- ggplot(charlie, aes(x = infected, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of Infected and Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(size = 12)) # Adjust title size

# Bar chart for "treatment" variable (percentage)
chart_percentage <- ggplot(df_percentage, aes(x = treatment, y = percentage, fill = treatment)) +
  geom_bar(position = "dodge", color = "black", stat = "identity") +
  labs(x = "Treatment", y = "Percentage") +
  ggtitle("Bar Chart of Treatment (Percentage)") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(size = 12)) # Adjust title size

# Combine charts side by side
library(gridExtra)
grid.arrange(chart_infected, chart_percentage, nrow = 1)


```


This study provides strong evidence of the drug's efficacy as demonstrated by the comparative infection rates among individuals who received the placebo and those who received the drug. Notably, the placebo group exhibited a significantly higher infection rate of approximately 80%, whereas the group receiving the drug had a considerably lower infection rate of approximately 25%. Moreover, when comparing the infection rates of individuals who did not contract AIDS despite taking either the drug or placebo, a clear distinction emerges, further supporting the drug's effectiveness. This is done through a visual analyse of the data using bar charts.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=charlie)
rowPerc(table1)
colPerc(table1)
```
Approximately 25% of individuals who contracted AIDS had taken the drug, while the remaining 75% had received the placebo. Among the non-infected group, approximately 97% had taken the drug, while 92% had received the placebo.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

There is a substantial statistical difference indicated by a p-value of .0000001242 in the numerical breakdown of the data regarding the use of the drug. This suggests that the drug has a significant impact on the outcomes being measured.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
Based on the collected data, the triUmpsal drug shows a clear efficacy among DrugUsers. Comparing the infection rates between the drug and placebo groups reveals a significant difference. Among the DrugUsers who took the drug, the infection rate was approximately 25%, while the placebo group had a considerably higher infection rate of around 80%.

Furthermore, when examining the non-infected individuals, a higher percentage (97%) of those who took the drug remained uninfected compared to those who received the placebo (92%). These results strongly indicate that the triUmpsal drug plays a vital role in reducing the incidence of HIV transmission among DrugUsers.

To enhance data analysis and visualization, bar charts can be utilized to provide a clearer depiction of the infection rates and highlight the notable disparities between the drug and placebo groups. These findings not only demonstrate the drug's efficacy but also suggest its potential to significantly contribute to the prevention of HIV transmission among DrugUsers.

Additionally, a p-value of .0000001242 reveals a substantial statistical difference because the number is so low it is something that we should pay attention too. In the numerical breakdown of the data, further supporting the conclusion that the drug has a significant impact on the measured outcomes.









